\section{V2 Objectives and Evolution from V1}

V2 is not a rewrite. The core architecture of V1 --- six independent microservices, database-per-service, API Gateway as the single entry point, Eureka for service discovery --- remains unchanged. V2 evolves the system within that architecture, addressing the gaps identified at the end of V1.

The V2 objectives were defined in five phases, ordered by impact and dependency:

\begin{enumerate}
    \item \textbf{Phase 1 --- Technical debt and correctness} --- Flyway migrations, filter corrections, MDC fix, role-based authorization, optimistic locking, and idempotency key. Correctness problems were resolved before adding new capabilities.
    \item \textbf{Phase 2 --- Kafka hybrid model} --- asynchronous replication of stable resource data from Catalog to Booking, reducing synchronous coupling while preserving real-time validation for availability.
    \item \textbf{Phase 3 --- Observability} --- Prometheus, Grafana, and Micrometer Tracing with Zipkin, replacing manual correlation ID logging with structured metrics and distributed traces.
    \item \textbf{Phase 4 --- Testing and CI/CD} --- integration tests with Testcontainers, consumer-driven contract tests with Pact, and a GitHub Actions pipeline.
    \item \textbf{Phase 5 --- Kubernetes} --- containerisation of all services, Kubernetes deployment manifests, and Flyway as the migration strategy for production schema management.
\end{enumerate}

\subsection{Why Correctness Comes First}

The decision to resolve correctness issues before introducing new capabilities was deliberate. Kafka, observability, and Kubernetes are additive --- they extend the system without changing its behaviour. Concurrency bugs, incorrect filter types, and missing authorization are subtractive --- they make the system behave incorrectly regardless of what is added on top.

Optimistic locking and the idempotency key were originally planned for Phase 4 alongside integration tests. They were moved to Phase 1 after recognising that concurrency is a correctness property, not a performance optimisation. A system that may create duplicate bookings under concurrent load is incorrect by definition, regardless of how well it is observed or how cleanly it is deployed.

\subsection{Continuity with V1}

V2 preserves all V1 design decisions that remain valid:

\begin{itemize}
    \item Database-per-service pattern --- each service owns its schema and data
    \item Stateless JWT authentication centralised at the API Gateway
    \item OpenFeign for synchronous inter-service communication
    \item Resilience4j retry, circuit breaker, and time limiter on Catalog calls
    \item RabbitMQ for centralised logging with correlation ID propagation
    \item Eureka for service discovery
\end{itemize}

Where V1 decisions are replaced or extended, this report explains the reasoning explicitly.